Individual Partisan Harm and Utility in Redistricting

### 2021-10-14 paper/paper.Rmd

---
title: Individual Partisan Harm and Utility in Redistricting
authors:
  - name: Cory McCartan
    thanks: |
      To whom correspondence should be addressed. 
      The authors thank ____ for helpful comments and discussion.
    department: Department of Statistics
    affiliation: Harvard University
    location: Cambridge, MA
    email: cmccartan@fas.harvard.edu
  - name: Christopher T. Kenny
    department: Department of Government
    affiliation: Harvard University
    location: Cambridge, MA
    email: christopherkenny@fas.harvard.edu
abstract: |
  Enter the text of your abstract here.
keywords:
  - partisan fairness
  - representation
  - redistricting simulation
  - individual standing
bibliography: references.bib
biblio-style: apalike
output:
  rticles::arxiv_article:
    keep_tex: true
    includes:
      in_header: "header.tex"
    latex_engine: pdflatex
---

@@

```{=html}
<!--
Districting is a complicated process. Residents of a state or locality must be partitioned into districts based on the geographies that they live in. These districts must adhere to many constraints, often requiring components such as contiguity and equal population. In drawing the districts, it's easy to lose sight of the individuals impacted by district drawing. Often, we consider how parties succeed or fail under the electoral environments generated by enacted plans. Yet this misses that parties are formed by voters who ally with parties. Rather than considering parties, we ask how individuals are harmed or benefited by the districting process?

Partisan symmetry is a useful quantity of interest, which has been widely used to improve academic and legal understanding of the redistricting process. We do not challenge the value of partisan symmetry, but instead suggest that a different question should be asked to describe the fairness of district based democracies.

Understanding fairness in electoral systems has long suffered from a focus on parties over the individuals who vote for them.
There is no doubt that parties are interesting units in electoral discussions.
They offer stability over candidates, information for the under-informed, and shortcuts for understanding preferred vote choices.
Yet, court histories and academic traditions value the harm of the individual over the harm to a transient political body, such as the party.

In this article, we offer an approach to understanding partisan fairness through a lens of individual harm and utility.
This does not supplant existing understandings of partisan fairness, but instead offers an alternative rooted in understanding how the individual potential voters are impacted by districting plans.

Several existing methods of redistricting simulation rely on randomness, such as Random Seed and Grow (Chen and Rodden 2013) or Recombination (Deford et al. 2021).
Moving beyond these heuristic methods, we can use sampling methods that have defined target distributions to ensure that we characterize the space of applicable plans properly.
Methods which operationalize a statistics-based approach include Fifield et al. 2020, McCartan and Imai (2020), and Autrey et al (2020).
Using these methods advance us beyond the critiques of Katz, King, and Roseblatt (2020) by sampling plans using weights within a Gibbs distribution, rather than assuming that uniform sampling weights alone are sufficient.

Metrics often associated with partisan symmetry, such as TODO, are indicative of violations of partisan symmetry.
Notably, these are only defined for full legislatures (see Katz, King, and Rosenblatt (2021)).
As such, theory-based interpretations of these measures do not indicate signs of unfairness to parities in the congressional district setting, unless they are consistent across many, many states.
We offer measures of partisan harm which are rooted directly in the representational harms inflicted upon individuals by those who choose to draw a map.

These measures are deeply connected to the law and constitutional understandings of partisan fairness, due to their related roots in the individual.

-   introduce partisan symmetry

    -   Katz, King, Rosenblatt 2019; Deford et al 2021; Katz, King, Rosenblatt 2021

-   Election law side is very indivudalized

    -   Standing discussions

        -   Lujan v Defenders of Wildlife (1992)
        -   Shaw v. Reno (1993) -- racial gerrymandering
        -   Issacharoff and Karlan 1998

-   Recent, related work

    -   EGap
    -   Partisan Dislocation

-   Important to consider benefits and harms to individual voters

-   Power of Ensemble Analysis

    -   Recent advances allow us to make claims about harm based in counterfactuals that are realistic and statistically useful
-->
```
## Running Examples

@@

We should say something about the importance in developing further measures which utilize simulation approaches

@@

\begin{proof}
\end{proof}
```
# Computing Utility, Harm, and Fairness from Precinct-level Election Data {#app:formulas}

@@

Formulas here.

# Districting Measures in New Jersey {#app:nj}

>>>>

---
title: Individual Partisan Harm and Utility in Redistricting
authors:
  - name: Cory McCartan
    thanks: |
      To whom correspondence should be addressed. 
      The authors thank {you} for helpful comments and discussion.
    department: Department of Statistics
    affiliation: Harvard University
    location: Cambridge, MA
    email: cmccartan@fas.harvard.edu
  - name: Christopher T. Kenny
    department: Department of Government
    affiliation: Harvard University
    location: Cambridge, MA
    email: christopherkenny@fas.harvard.edu
abstract: |
  Most existing standards for assessing partisan gerrymandering focus on the symmetric treatment of parties by electoral systems. We develop measures of utility and harm for individual voters by considering counterfactual redistricting plans, and use these notions to construct a normative standard of partisan fairness which is centered on the individual. We estimate these measures from real-world data by leveraging recent advances in redistricting sampling methodology, which allow researchers to generate large numbers of districting plans which reflect local legal and practical requirements. We apply our measures to congressional redistricting in New Jersey, and compare them to existing measures of partisan symmetry and competitiveness.
keywords:
  - partisan fairness
  - representation
  - redistricting simulation
  - individual standing
bibliography: references.bib
biblio-style: apalike
output:
  rticles::arxiv_article:
    keep_tex: true
    includes:
      in_header: "header.tex"
    latex_engine: pdflatex
editor_options: 
  markdown: 
    wrap: sentence
---

@@

## Running Examples

@@

This paper is an attempt to bridge the quantitative gap between votes and representation.
We break from existing redistricting literature and focus on the relationship between votes and representation by copartisans, rather than the relationship between votes and seats.
This focus naturally leads to a new normative standard of partisan fairness, which is achieved when the necessary harms of a districting plan are distributed evenly across voters of each party.
This may occur regardless of whether the districting plan achieves partisan symmetry.

This approach offers new purchase into the foundational questions of how to quantify the harms and benefits that accrue to voters in district-based democracies, where the rules of defining districts are fixed, but the choice of how to how to draw the districts remains a policy question.
While the overall framework is flexible, we offer a selection of simple yet interesting measures for improving the individual partisan implications of a districting plan.
Further work is needed to characterize how existing redistricting requirements and practices, such as the number or size of districts, or encouragement of competitive districts, interact with these measures.

@@

\begin{proof}
\end{proof}
```
```{=html}
<!--
# Computing Utility, Harm, and Fairness from Precinct-level Election Data {#app:formulas}

@@

Formulas here.
-->
```
# Districting Measures in New Jersey {#app:nj}
